---
title: "NYPD Activities"
author: "Anne-Gaelle Songeons"
output: 
      rmdformats::downcute:
        code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r library, message=FALSE, warning=FALSE, include=FALSE}
library(viridis)
library(tidyverse)
library(pastecs)
library(leaflet)
library(ggplot2)
library(forcats)
library(tigris)
```


```{r data, message=FALSE, warning=FALSE, include=TRUE}

#LOADING THE DATA
file_path <- file.path("Data","NYPD_Arrest_Data__Year_to_Date_.csv") #To make sure the code is going to run for every system 
file_path

NYPD_Arrest <- read.csv(file = file_path, stringsAsFactors = FALSE) #if the user is using an older version of r 

```

```{r data2, message=FALSE, warning=FALSE, include=TRUE}
#load map
file_path2 <- file.path("Data","geo_export_973a007b-36dc-4dfb-b474-58657b7086be.shp")
ny_map <- sf::st_read(dsn = file_path2)
```

# Research Question 

On December 14, 2021, future New York City Mayor Eric Adams nominated **the first woman and third person of color to head the New York City Police Department: Chief Keechant Sewell**. This decision is historic, especially in a tense American context where many have lost confidence in the police. 
Often accused of being **poorly trained, of racism and police violence** - we think here of the death of George Floyd in 2020 that shook the United States and the world -, it is a great challenge that awaits the new chief. 

We propose in this report to make a point on the **activities of the police in the city of New York**. 
For this we will use the data provided by the city of New York. We found a database containing all the arrests of the year 2021 (from January to October) which is available [here](https://data.cityofnewyork.us/Public-Safety/NYPD-Arrest-Data-Year-to-Date-/uip8-fykc). 


<br>  

![](https://images.unsplash.com/photo-1493603500863-d1f8f3b2386d?ixlib=rb-1.2.1&ixid=MnwxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8&auto=format&fit=crop&w=2340&q=80)



<p style="font-size: 24pt"> **What is the state of NYPD arrests in 2021?** </p>

To answer our problem we will analyse into consideration two aspects: 

- The daily activities of the policemen
- The profile of the arrested people 

# Data Description 


```{r eval=FALSE, include=FALSE}
pastecs::stat.desc(NYPD_Arrest) #display main stats about our data
```


```{r cleaning, include=FALSE}
NYPD_Arrest_Clean  <-  na.omit(NYPD_Arrest) # drop all the rows with at least 1 NA
  
NYPD_Arrest_Clean  <-  unique(NYPD_Arrest_Clean) # check if there is duplicate

NYPD_Arrest_Clean <-  dplyr::select(NYPD_Arrest_Clean,
                                    ARREST_DATE, 
                                    OFNS_DESC, 
                                    LAW_CAT_CD, 
                                    ARREST_BORO,
                                    AGE_GROUP, 
                                    PERP_SEX, 
                                    PERP_RACE, 
                                    Latitude, 
                                    Longitude)
#Select the row we are interested in 


#to change the values in the Arrest_Borro column, it will simplify our future analyses
NYPD_Arrest_Clean$ARREST_BORO[NYPD_Arrest_Clean$ARREST_BORO == 'B'] = 'Bronx'
NYPD_Arrest_Clean$ARREST_BORO[NYPD_Arrest_Clean$ARREST_BORO == 'K'] = 'Brooklyn'
NYPD_Arrest_Clean$ARREST_BORO[NYPD_Arrest_Clean$ARREST_BORO == 'M'] = 'Manhattan'
NYPD_Arrest_Clean$ARREST_BORO[NYPD_Arrest_Clean$ARREST_BORO == 'Q'] = 'Queens'
NYPD_Arrest_Clean$ARREST_BORO[NYPD_Arrest_Clean$ARREST_BORO == 'S'] = 'Staten Island'



                     
```

Our data didn't need much cleaning, we just detected a **problem with missing values**. We decided to delete the rows that had at least one missing value. So we go from `r nrow(NYPD_Arrest)` rows to `r nrow(NYPD_Arrest_Clean)` rows. 

We have selected the columns that will be useful for our analysis to simplify our dataset and change the values in the *ARREST_BORO* column. 




```{r echo=FALSE, fig.width=4}

#display as a table the 5 first rows of our dataset
knitr::kable(head(NYPD_Arrest_Clean, 5), caption = "Data after cleaning")

```



# NYPD Daily Activities 


## Over the year 

```{r echo=FALSE, fig.align="center", fig.height=8, fig.width=8}

# Reformart the date, and keep only the month 
NYPD_Arrest_Clean$MONTH <- format(as.Date(NYPD_Arrest_Clean$ARREST_DATE, format = "%m/%d/%Y"), "%m")


# Create a subset 
NYPD_Arrest_bydate  <-  NYPD_Arrest_Clean %>% 
                        group_by(MONTH) %>% 
                        summarise(Count = n())

p <- ggplot(NYPD_Arrest_bydate, aes(x=MONTH, y=Count)) + 
  #We specify here that we want to obtain a lollipop graph
    geom_segment( aes(x=MONTH, xend=MONTH, y=10000, yend=Count), color="grey") +
  #We choose the size and the color of the point
    geom_point(size=20, color="#2B8F59") + 
  #We reverse the x-axis and y-axis to make the graph more readable
    scale_x_discrete(labels=c("JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP"))+
  ylim(10000,14000) +
  #We insert the values at the end of the points
    geom_text(aes(label=Count), nudge_y = 3, color="white", size=3) +
  #We choose the title of the graph
    labs(title = "Count of arrests per month in 2021", size= 3)+
  #We choose the aesthetic of the graph and remove the names of the axes to avoid overloading 
  theme_minimal() +
  theme(
    axis.title =element_blank(),
    panel.grid = element_blank(),
    legend.position = "none", 
    axis.text.y = element_blank())
    

p
```



On average, the New York Police  make  **`r as.integer(mean(NYPD_Arrest_bydate$Count))` arrests ** per month over the course of 2021. 
We note that in April the number of arrests is well below this average, the police made `r NYPD_Arrest_bydate[NYPD_Arrest_bydate$MONTH == "04", ][2]` arrests. It is the month where the least number of arrests was recorded. The possibility that this is related to the Covid health crisis is not to be dismissed.



## By Borough

```{r echo=FALSE, message=FALSE, warning=FALSE}

#Create a subset
NYPD_Arrest_monthboro  <-  NYPD_Arrest_Clean %>% 
                        group_by(ARREST_BORO, MONTH) %>%  
                        summarise(Count = n()) %>% 
                        arrange(Count) %>% 
                        mutate(ARREST_BORO = fct_reorder(ARREST_BORO, Count))
        

q <- 
  ggplot(NYPD_Arrest_monthboro, aes(x = reorder(ARREST_BORO, Count), y =  Count)) +  #plot the data as a boxplot
  geom_boxplot(aes(fill=ARREST_BORO), color="#37328c", size=0.5)  +
  
  #add color on top of it
  scale_fill_viridis(discrete = T, direction=-1, guide = guide_legend(reverse = TRUE))+
  
  #Change the orientation of the graph
  coord_flip()+
  
  scale_x_discrete(labels=c("Staten Island", "Queens",  "Bronx", "Manhattan", "Brooklyn")) + 
  
  #Custom the graph to make it clearer
  theme_minimal() +
    theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour = "white"),
        legend.position = "none")+
  labs(x = "", 
       y = "Number of arrests per month (mean)", 
       title = "Distribution of arrests by borough per month for 2021"
  )

q

```

The following graph represents the average number of crimes per month and per neighborhood in New York. There is a **great disparity** between the neighborhoods of New York in terms of crime. 
**Staten Island is very little concerned** where Brooklyn comes first. 

This can be correlated to the **very different standards of living** in the different parts of the city. Staten Island clearly stands out as the richest of the 5 boroughs and the least populated. 


```{r echo=FALSE, message=FALSE, warning=FALSE}


#subset
NYPD_Arrest_byboro <- NYPD_Arrest_Clean %>%  
                        group_by(ARREST_BORO) %>% 
                        summarise(Count = n()) 

# Merge data
boro_merged <- geo_join(ny_map, NYPD_Arrest_byboro , "boro_name", "ARREST_BORO")

# color palette + popup 
pal <- colorNumeric("Blues", domain=boro_merged$Count)
popup_sb <- paste0("Borough: ", as.character(boro_merged$boro_name), "</br>", 
                   "Total: ", as.character(boro_merged$Count))

# plot with leaflet
leaflet(ny_map) %>%
  addPolygons(
    fillColor = ~pal(boro_merged$Count),
    fillOpacity = 1, 
    stroke = FALSE, 
    color = 'White',
    weight = 0.2, 
    popup = ~popup_sb
    ) %>%
  addLegend(pal = pal, 
            values =  boro_merged$Count, 
            position = "bottomright", 
            title = "Total number of arrest in 2021")

```

<style>
.leaflet-container {
    background: #F6F5F5!important;
}
</style>

## Crime and offence

```{r message=FALSE, warning=FALSE, include=FALSE}

#Create a subset
NYPD_Arrest_byoffense0  <- NYPD_Arrest_Clean %>% 
                        group_by(OFNS_DESC) %>% 
                        summarise(Count = n()) 
                       


# select the offences which are less represented in the dataset
minorOffense = NYPD_Arrest_byoffense0$OFNS_DESC[NYPD_Arrest_byoffense0$Count < 1000] 

# replace those offenses by "Other"
# important when we will plot the graph to avoid having to much data
NYPD_Arrest_Clean[ NYPD_Arrest_Clean$OFNS_DESC  %in% minorOffense,  ]$OFNS_DESC = "OTHERS"


# Create the subset we will use for the next graph
NYPD_Arrest_byoffense  <- NYPD_Arrest_Clean %>%  
                          group_by(OFNS_DESC) %>% 
                          summarise(Count = n())  %>% 
                          arrange(Count) %>% 
                          mutate(OFNS_DESC=factor(OFNS_DESC, OFNS_DESC))

```



```{r echo=FALSE, fig.height=11, fig.width=8, message=FALSE, warning=FALSE, fig.align="center"}
# create a df for the label
NYPD_Arrest_byoffense$id = seq(1,nrow(NYPD_Arrest_byoffense))
label_data <- NYPD_Arrest_byoffense 
 
# calculate the angles for the labels
number_of_bar <- nrow(label_data)
angle <-  90 - 360 * (label_data$id-0.5)/number_of_bar     
# calculate the alignment of labels
label_data$hjust<-ifelse( angle < -90, 1, 0)
# flip label angle 
label_data$angle<-ifelse(angle < -90, angle+180, angle)

#plot the graph
k <- ggplot(NYPD_Arrest_byoffense ) + 
  geom_bar(aes(x = as.factor(id), y = Count, fill = OFNS_DESC), stat = "identity")  +  
  
  # add colors
  scale_fill_viridis(discrete=T, direction = -1)+
  
  # make the plot circular
  coord_polar() +
  
  # define the size of the circle
  ylim(-10000,40000)+ 
  
  #add customization 
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-3,4), "cm"), 
    legend.position = "none"
    ) + 
  
  # add the labels
  geom_text(data=label_data, aes(x=as.factor(id), y=Count+200, label=OFNS_DESC, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=3, angle= label_data$angle, inherit.aes = F )  +
  
  # add a title
  geom_text( aes(x=22, y=-10000, label="Offenses in NYC"), color="black", inherit.aes = FALSE, size = 4) 
 
k

```


The illegal acts perpetrated are very diverse, we count not **far from `r nrow(NYPD_Arrest_byoffense0)` categories** on this year. 
Police officers are **most commonly confronted with third degree assaults and related offenses**. This includes, for example, violent attacks on law enforcement, firefighters, and medical personnel. This phenomenon **represents  `r round(NYPD_Arrest_byoffense$Count[NYPD_Arrest_byoffense$OFNS_DESC == "ASSAULT 3 & RELATED OFFENSES"] * 100/sum(NYPD_Arrest_byoffense$Count), 2)`% of arrests**. 


**Where the arrests for assault of third degree took place? **
</br>

```{r echo=FALSE, message=FALSE, warning=FALSE}
#subset
Assault3 <- NYPD_Arrest_Clean %>% 
  filter(OFNS_DESC == "ASSAULT 3 & RELATED OFFENSES") 

#plot with leaflet
leaflet( Assault3 ) %>% 
  addTiles() %>% 
  addMarkers(lng = Assault3$Longitude , 
             lat = Assault3$Latitude, 
             popup = Assault3$ARREST_DATE, 
             clusterOptions = markerClusterOptions()) 

```




# Profile of the arrested people   



## Age & Sex
```{r echo=FALSE, message=FALSE, warning=FALSE}

# Subset the data
NYPD_Arrest_byagesex <-  NYPD_Arrest_Clean %>% 
                        group_by(AGE_GROUP,PERP_SEX) %>% 
                        summarise(Count = n()) 

# assign negative values to female in order to have a "miror" of the male 
NYPD_Arrest_byagesex$Count[NYPD_Arrest_byagesex$PERP_SEX == "F"] = -1*NYPD_Arrest_byagesex$Count

```


```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
x <- ggplot() +
  
  # plot the data for the male
  geom_bar(data = filter(NYPD_Arrest_byagesex, PERP_SEX == 'M'), aes(x = AGE_GROUP,  y = Count, fill = PERP_SEX, alpha = 0.8), stat = "identity") +
  
  #plot the data for the female 
  geom_bar(data = filter(NYPD_Arrest_byagesex, PERP_SEX == 'F'), aes(x = AGE_GROUP,  y = Count, fill = PERP_SEX), stat = "identity") +
  
  #change the orientation
  coord_flip() +
  
  # customize the graph
  theme_minimal() +
  theme (
    axis.title = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position = "none") +
  labs(y = "Count", x = "Age Band", title = " Age distribution of arrested citizens by gender  ") +
  scale_fill_manual(values = c("#440154", "#28ae80"))+
  ylim( -30000, 60000) 
  
x

```

Here two things are to be noted: 
- the **majority of criminal acts are perpetrated by men**
- And the **most affected age group is the 25/45 years old** 

These are the publics with which the police are most in contact. This is interesting information, because it can allow the police to be more attentive to certain types of people but also in order to **recreate links to know who to communicate with**. 

## Race 


```{r echo=FALSE, fig.align="center", fig.height=8, fig.width=8, message=FALSE, warning=FALSE}

# Subset
NYPD_Arrest_byrace <-  NYPD_Arrest_Clean %>% 
                        group_by(PERP_RACE) %>% 
                        summarise(Count = n()) %>%
                        arrange(desc(Count)) %>% 
                        mutate(PERP_RACE=factor(PERP_RACE, PERP_RACE))

# create a new column with the percentage 
NYPD_Arrest_byrace$perc <- NYPD_Arrest_byrace$Count/sum(NYPD_Arrest_byrace$Count)


m <- ggplot(data = NYPD_Arrest_byrace, aes(x = PERP_RACE, y=perc))  + 
  
  
  geom_col(aes(fill = perc == max(NYPD_Arrest_byrace$perc))) +  # here we want to have only one column with a specific color 
  
  # customization 
  scale_fill_manual(values = c('grey', '#440154'))+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+ 
  theme_minimal() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_blank(),
        axis.title = element_blank(), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), 
        legend.position = "none")+
  labs(
       title = "Percentage of arrested in 2021 according to their race"
  ) 
  

m

```

It is noted that nearly **half of the people** who are arrested in New York City are people **from the Black American community**. This is all the more alarming considering that in 2020 New York City reported only 15 and 20% of people from this community. 
There is therefore an overrepresentation of these people. 

We can put forward several hypotheses that can explain these figures, such as **poverty and difficult social environments in this community**, which would lead them to commit more crimes later on, but also we take the **discriminations** of the police towards these people in particular. 
These hypotheses remain to be verified with more robust data. 


```{r echo=FALSE, fig.height=10, fig.width=8, message=FALSE, warning=FALSE, fig.align="center"}

#subset 

NYPD_Arrest_byraceboro <-  NYPD_Arrest_Clean %>% 
                        group_by(ARREST_BORO, PERP_RACE ) %>% 
                        summarise(Count = n()) %>%
                        transmute(PERP_RACE , perc = Count/sum(Count), max = max(perc)) #Add a column with the maximum per neighborhood 
   

m <- ggplot(data = NYPD_Arrest_byraceboro, aes(x = reorder(PERP_RACE, - perc), y=perc))  + # to order the graph 
  
  geom_col(aes(fill = perc == NYPD_Arrest_byraceboro$max)) + # choose which column to color, the maximun here 
  
  #customization
  scale_fill_manual(values = c('grey', '#440154'))+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+  #add percentages to the scale 
  theme_minimal() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_blank(),
        axis.title = element_blank(), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), 
        legend.position = "none")+
  labs(
       title = "Percentage of arrested in 2021 according to their race"
  ) + 
  
  # display a graph for each neighborhood
  facet_wrap( ~ ARREST_BORO, ncol = 1)

m

```


What we have demonstrated above applies to all areas of the city. **African American persons are overrepresented**. 


# Conclusion 


The challenges of policing in New York are great. We have highlighted in this report that there is a **great disparity between New York neighborhoods**, and **alleged differences in treatment according to ethnicity**. 
This is symptomatic of an America that is more than ever divided on all fronts. 

The police, being part of the daily life of its citizens, has here its role to play to make decrease the crime rate in the city. They can indeed play a role of mediator, and recreate social links.  

